The watershed transform in ITK - discussion and new developments

### 2006-04-25 article/Article.tex
pronucleus segmentation in bovine embryo
darcs-hash:20060425150425-2fc9d-474338ada516754e8a5ad079e60e8367a9501286.gz

\caption{Three difficult test images. The first image, plateau, show that ITK's watershed produce a bad result on this image: only one minima take all the image. This problem is also shown in Figure \ref{fig:cthead1-markers}. The morphological watershed give a better result, but not yet perfect: the separation line should not be curved as is it. The next image, thick line, again show that ITK watershed can produce strange results with a green label which is where we should have the light blue one. Morphological watershed gives good results. The last image, button hole, is a very difficult case for watershed algorithms which mark watershed line: the watershed pixel at the bottom of the image stop the label propagation. There is several possible behavior: produce a large watershed line which cover all the zone (Vincent-Soille algorithm, not shown here), let the label on the top of the image propagate in the zone (Meyer algorithm), split the zone in 2 zones with the 2 labels of the bottom of the image (topological watershed, not shown here).\label{fig:difficult-cases}}
\end{center}
\end{figure}

@@

>>>>

\caption{Three difficult test images. The first image, plateau, show that ITK's watershed produce a bad result on this image: only one minima take all the image. This problem is also shown in Figure \ref{fig:cthead1-markers}. The morphological watershed give a better result, but not yet perfect: the separation line should not be curved as it is. This problem is also visible in Figure \ref{fig:embryo-mws}. The next image, thick line, again show that ITK watershed can produce strange results with a green label which is where we should have the light blue one. Morphological watershed gives good results. The last image, button hole, is a very difficult case for watershed algorithms which mark watershed line: the watershed pixel at the bottom of the image stop the label propagation. There is several possible behavior: produce a large watershed line which cover all the zone (Vincent-Soille algorithm, not shown here), let the label on the top of the image propagate in the zone (Meyer algorithm), split the zone in 2 zones with the 2 labels of the bottom of the image (topological watershed, not shown here).\label{fig:difficult-cases}}
\end{center}
\end{figure}

@@

In this example, we use the watershed filter to split two nearly round
objects: the male and female pronucleus in a bovine embryo, around 18
hours after fecondation.

The program is in python for conciseness\footnote{And also to promote
python wrappers.}.

\begin{figure}[htbp]
\begin{center}
\subfigure[Input]{\includegraphics[scale=0.22]{embryo}}
\subfigure[median]{\includegraphics[scale=0.22]{embryo-median}}
\subfigure[threshold]{\includegraphics[scale=0.22]{embryo-th}}
\subfigure[opening by reconstruction]{\includegraphics[scale=0.22]{embryo-open}}
\subfigure[fill hole]{\includegraphics[scale=0.22]{embryo-fill}}
\subfigure[invert]{\includegraphics[scale=0.22]{embryo-invert}}
\subfigure[distance]{\includegraphics[scale=0.22]{embryo-distance}}
\subfigure[inverted distance]{\includegraphics[scale=0.22]{embryo-idistance}}
\subfigure[morphological watershed\label{fig:embryo-mws}]{\includegraphics[scale=0.45]{embryo-overlay}}
\subfigure[ITK watershed]{\includegraphics[scale=0.45]{embryo-overlay2}}

\caption{The results of the different steps of the segmentation of bovine pronucleus.}
\end{center}
\end{figure}

\small \begin{verbatim}
import itk
reader = itk.ImageFileReader.IUC2.New(FileName='embryo.png')
\end{verbatim} \normalsize

We use a median image filter to remove some noise, and threshold the image
to get a binary image\footnote{The methods to remove the noise and binarize the
image are volontary kept simple for this example}.

\small \begin{verbatim}
median = itk.MedianImageFilter.IUC2IUC2.New(reader)
th = itk.BinaryThresholdImageFilter.IUC2IUC2.New(median, LowerThreshold=40)
\end{verbatim} \normalsize

The binary image produced by the \verb$ThresholdImageFilter$ contains lot
of small objects. To remove them, we use an {\em opening by reconstruction}
which have the advantage to preserve the egdes of the object.

\small \begin{verbatim}
kernel = itk.strel(2, 30)
open = itk.OpeningByReconstructionImageFilter.IUC2IUC2.New(th, Kernel=kernel)
\end{verbatim} \normalsize

There is now only one binary object in the image, but it contains lots of holes
which will produce an unsable distance map for the watershed. We fill them
with a {\em hole filling} filter. This algorithm, based on a morphological
reconstruction, fill the hole in a binary object but preserve the edges of the
object. In this case, it is important to keep the two cavities between the two
pronucleus unchanged.

\small \begin{verbatim}
fill = itk.GrayscaleFillholeImageFilter.IUC2IUC2.New(open)
\end{verbatim} \normalsize

Then we compute the distance map which will be used by the watershed. The
binary object must be inverted to compute the distance {\em inside} the
object, and the distance map must be inverted to used by the watershed.

\small \begin{verbatim}
invert = itk.InvertIntensityImageFilter.IUC2IUC2.New(fill)
distance = itk.DanielssonDistanceMapImageFilter.IUC2IUC2.New(invert)
idistance = itk.InvertIntensityImageFilter.IUC2IUC2.New(distance)
\end{verbatim} \normalsize

The watershed is then performed to split the two pronucleus. A threshold
is used to avoid the too small minima in the border of the object and thus
avoid over segmentation. The label produced by the watershed are then
displayed on top of the original image.

The watershed extends the labels on the entire image. We use a
\verb$MaskImageFilter$ to limit the labels to the object. Another commonly
used method is to use the background as a marker and to perform a watershed
from markers.

\small \begin{verbatim}
ws = itk.MorphologicalWatershedImageFilter.IUC2IUC2.New(idistance,
                                                        MarkWatershed=False,
                                                        Threshold=10)
mask = itk.MaskImageFilter.IUC2IUC2IUC2.New(ws, fill)
overlay = itk.LabelOverlayImageFilter.IUC2IUC2IRGBUC2.New(reader, mask, Opacity=0.5)
\end{verbatim} \normalsize

The same watershed is done with the ITK watershed filter\footnote{A
$RelabelComponentImageFilter$ is used here because unsigned long is
not available in python for the $MaskImageFilter$}.

\small \begin{verbatim}
ws2 = itk.WatershedImageFilter.IUC2.New(idistance, Level=0.5)
relabel = itk.RelabelComponentImageFilter.IUL2IUC2.New(ws2)
mask2 = itk.MaskImageFilter.IUC2IUC2IUC2.New(relabel, fill)
overlay2 = itk.LabelOverlayImageFilter.IUC2IUC2IRGBUC2.New(reader, mask2, Opacity=0.5)
\end{verbatim} \normalsize

ITK watershed clearly produce a wrong segmentation in that case.
The result of the morphological watershed is quite good, even if not perfect:
there is some curves in the separation line of the two labels. This is the
result of the distance approximation in the watershed algorithm.
